\mytitle

### 2015-08-02 MSc_diss.tex
Rewriting literature review

\subsection{Ozone, alkyl nitrates and carbon bonds}

@@

@@

\subsubsection*{Lifetime}
\citep{Newland2013}
[Species within this group provide a transport reservoir for the short lived NOx compounds due to their relatively longer lifetimes. This allows NOx to be transported from polluted areas to remote background sites. Alkyl nitrates have been reported as making up only a few percent of the NOy ('odd' nitrogen) budget in urban and rural areas (Shepson et al., 1993; Thornberry et al., 2001), however due to their relatively long lifetimes and their continued production during transport of polluted air masses, as alkanes are converted to alkyl nitrates (e.g. Bertman et al., 1995), they can comprise up to tens of percent of the NOy budget in polar regions and in the remote marine troposphere (Talbot et al., 2000). Therefore alkyl nitrates could potentially be a substantial source of NOx in remote background regions contributing to tropospheric ozone formation in these regions.]
\citep{Newland2013}
[The lifetimes of the alkyl nitrates vary with latitude and season due to the seasonality of the OH sink and photolysis. Clemitshaw et al. (1997) report a range of lifetimes with respect to OH and to photolysis. Photolysis lifetimes vary from about ten days for butyl and pentyl nitrate in the summer to 1 – 2 years in the winter at 65N with lifetimes also decreasing with altitude. The lifetimes due to reaction with OH (tauOH) at 65N range from tens of years in
winter to about 10 days in the summer. tauOH increases with altitude as the OH concentration decreases. At the equator tauOH is similar all year round, on the order of a few days.]
\citep{Chuck2002}
Use of the mean sea-to-air fluxes for 0 to 40S (Table 1) and measured atmospheric concentrations results in a lifetime of be- tween 4.5 and 25 days for MeONO2 and between 5 and 10 days for EtONO2 (22). Literature estimates of the atmospheric life- time of MeONO2, based on known atmospheric destruction processes, are poorly con- strained, ranging from 6 to 29 days (23, 24). For EtONO2
, the atmospheric lifetime with respect to photodissociation at the equator and at 40S has been estimated to be 7 and 12 days, respectively, for October (25).

@@

\subsubsection{Comparison methodology}
\citep{Sommariva2008}
[The ratios between the alkyl nitrates and their
main alkane precursors can also be plotted against one another, for example against 2-butylnitrate/ n-butane as in Fig. 4. In this way the modelled and the measured concentrations can be compared without taking into account their variation with time. A similar approach was used in previous studies to investigate the sources of alkyl nitrates and the photochemical processing of polluted air masses (e.g., Bertman et al., 1995; Flocke et al., 1998; Roberts et al., 1998; Stroud et al., 2001; Simpson et al., 2003; Reeves et al., 2007). These studies compared plots of the measured ratios against 2-butylnitrate/n-butane to the ratios calcu- lated with a simple sequential reaction model which assumed a single precursors for each alkyl nitrate. Usually, the underestimation of the ratios for short- chain compounds (propylnitrates/propane, ethylni- trate/ethane) was interpreted as an indication of photochemical sources other than the main alkane precursor.]

@@

\citep{Atherton1989}
The transport from urban areas can lead to alkyl nitrate concentrations of 4-83 pptv after 120- 240 hours. These levels are far higher than those possible from marine in-situ production. Additionally, alkyl nitrates may be more abundant than PAN in remote marine environments.
Model: box
Chemistry: explicit chemical reaction mechanism
[Carter et al., 1986. A variety of alkyl nitrates, including methyl-, butyl-, pentyl-, hexyl-, heptyl-, octyl-, and nonyl-nitrates are allowed to form from fourteen different alkyl peroxy radicals. + ocean emission sources fluxes of c
for ethene, ethane, propene, and propane, NO, NO2
Initial conditions: clear sky conditions, a quantum yield of 1, solar zenith angle of 45 degrees, and total ozone column of 8*10 cm-2. Daytime+nigthtime, invertion height, dilution. 298 K. No deposition.
Scenarios: Two, MLOW and MHIGH, represent the marine in-situ production of alkyl nitrates. The total hydrocarbon concentration was 17.6 ppbC. MLOW had NO = 4 pptv and NO2 = 4.6 pptv based on measurements over the Pacific Ocean. MHIGH had NO = 46 pptv, NO2 = 54 pptv, and PAN = 50 pptv. The two other scenarios, UHIGH and ULOW, simulate the transport of species from urban to remote environments. UHIGH contained 2150 ppbC and 140 ppbv NOx, and ULOW contained 270 ppbC and 18 ppbv NOx, giving each a HC/NOx ratio of 15.
Simulation time: 12 hours MLOW and MHIGH, 240 hours UHIGH and ULOW.

Results:
- The total alkyl nitrate concentration present after 12 hours is less than 0.4 pptv for MHIGH and less than 0.1 pptv for MLOW. These levels are much lower than observations, indicating that marine in-situ production can account for only a fraction of the total marine alkyl nitrate concentration.
- IHIGH The total alkyl nitrate concentration peaks during the first 6 hours and then drops to 3.1 ppbv after 12 hours. This value is com- parable with the 2 ppbv alkyl nitrate concentration de- termined in Los Angeles [Hanst et al., 1982]
- The model predicts alkyl nitrate concentrations of 21-83 pptv for UHIGH and 4-12 pptv for ULOW for the hours 120-240.
- General: The model predictions of 12-83 pptv (af- ter 120 hours) and 4-21 pptv (after 240 hours) are of the same order as observations.

\citep{Sommariva2008}
Validation: typical urban plume in the North-East of the United States, measurements taken during three flights of the NOAA WP-3D aircraft, which sampled a plume from the New York City area during the New England Air Quality Study (NEAQS) 2004 campaign
- Seven alkyl nitrates (ethyl- nitrate, i-propylnitrate, n-propylnitrate, 2-butylni- trate, 2-pentylnitrate, 3-pentylnitrate, 3-methyl-2- butylnitrate) were measured during these flights together with a full suite of VOCs and other chemical parameters, such as CO, O3,NOx, HNO3 and PANs (Neuman et al., 2006).
- The model results have been qualitatively compared to measurements of alkyl nitrates taken on 20–22 July 2004, when the NOAA WP-3D aircraft made three flights aimed to inter- cept the New York City plume.
Model: subset! of MCM v3.1 (The MCM is one of the most detailed chemical mechanism for tropospheric chemistry available)
- to study the formation processes and photochemical evolution of alkyl nitrates
Chemistry: oxidation mechanisms of 64 VOCs (50 hydrocar-
bons and 14 oxygenates) plus CH4 and CO, The inorganic chemistry mechanism was taken from the IUPAC recommendation (Atkinson et al., 2003). A dimethyl sulphide (DMS) mechanism (Sommariva et al., 2004) was also included. The photolysis rates were calculated using the MCM parametrization (Saunders et al., 2003) for end of July 2004 at the latitude and longitude of Boston, MA. only gas-phase!!! Measurements made during NEAQS 2004 showed that there was very little removal of, for example,HNO3 by dry deposition in plumes decoupled from the marine boundary layer during transport from New York City (NYC) to the North Atlantic (Neuman et al., 2006). Therefore, removal by dry deposition was neglected in the model.
Initialisation: measurements taken in the Boston metropolitan area during the 2002 and 2004 NEAQS campaigns. The initial concentrations of VOCs were determined using emission ratios; CH4 and O3 were set to typical values for urban conditions (1800 and 90 ppb) (Table 1). The model has been initialized using measurements and VOCs emission ratios from the 2002 and 2004 New England Air Quality Studies.
Simulation time: 5 days.
Scenarios: 2: A (higher OH, good with obs, reference), B
Time: 20-22 July.

Results:
- While long-chain (C5) alkyl nitrates are produced for 90\% or more from the oxidation of a single parent alkane, short-chain <C4 alkyl nitrates can be formed from several precursors. The relative importance of each production route has been quantitatively determined thanks to the high level of chemical detail provided by the MCM. These secondary routes
to the formation of alkyl nitrates include the oxidation of longer-chain alkanes and oxygenated intermediates, like carbonyls, peroxides and carboxylic acids.
- The best agreement is in the case of 3-methyl-2-butylnitrate and the discrepancy between the modelled and the measured ratios clearly increases with decreasing carbon number .
- RH/AN, Except in the case of ethylnitrate, the model can reproduce the measured rate of increase in the alkyl nitrate to alkane ratios fairly well, which suggests that the kinetic parameters used in the MCMmodel are overall correct.
- By considering together Figs. 3 and 4 it can be reasonably concluded that the chemistry of alkyl nitrates is well represented in the MCM model.
- Using the explicit description of the chemistry provided by the MCM, it has been possible to determine quantitatively the photochemical formation and the evolution of alkyl nitrates and the relative importance of their different formation pathways. The results are summarized in Table 2. Large alkyl nitrates are produced almost exclusively from the reaction of a single parent
alkane. Smaller <=C4 alkyl nitrates also have other significant precursors, like longer-chain alkanes and alkenes. For n-propylnitrate and ethylnitrate, pro- panal and butanal, which are largely emitted, are important sources within the first 24 h. For all alkyl nitrates, the role of organic peroxides and car- boxylic acids formed by the reactions of peroxy radicals with HO2 at low [NOx] increases with time. These species contribute to sustain the production of alkyl nitrates far from the emission points of the parent VOCs. The relative importance of the different pathways
and how they change with time are dependent on the initial conditions of the model. In a scenario with lower O3 and CO and higher NOx the
importance of the direct routes (alkane ?OH) is typically lower during the first two–three days, and the contributions of peroxides and carboxylic acids become significant later and are on average less important. These results suggest that secondary chemistry in the formation of the alkyl nitrates, particularly oC4, cannot be neglected and should be included in the chemical models, due to its potential impact on ozone-controlling strategies.

\citep{Newland2013}
Model1: 2-D atmospheric model
- to study the formation, loss and transport of the alkyl nitrates to the Arctic
- to derive northern hemisphere emissions of the six alkanes.
Chemistry:
- (with fixed NOx and OH field taken from full chemistry run with model by Hough and Johnson (1990))
Validation of ANs: comparison with measurements of alkyl nitrates in the Arctic.
Performance: recreate both the measured seasonality of the alkyl nitrate, and the peak mixing ratio, reasonably well.
Simulation time: 6 days?

Model2: box
Chemistry:
- alkanesmethane, ethane, propane, butane and pentane and ANs explicitly described (RO+NO2 too)

Model3: firn model
- to derive atmospheric histories of six alkanes and six alkyl nitrates from measurements in firn air from the NEEM site, Greenland (methyl, ethyl, 2-propyl, 2-butyl, 2+3-pentyl and 3- methyl-2-butyl nitrate), 1950-2008.

Results:
- There are two ways in which NOx mixing ratios can limit alkyl nitrate production: either through air masses being NO limited with respect to the formation of a given alkyl nitrate or by the NOx in an air mass being 'used up' (track back to Reeves)
-- Alkyl nitrate production becomes NOx saturated above a given NO mixing ratio. This limit increases with increasing carbon number.
-- The time taken for an air mass to 'use up' all the initial NOx is dependent mainly on the initial NOx mixing ratio.
- At high [NO2] (> 27 ppb), methyl nitrate production is dominated by the reaction of the alkoxy radical with NO2
- The regions that contribute most to modeled alkane and alkyl nitrate mixing ratios in the Arctic are Europe and Russia
- The importance of the RO + NO2 pathway compared to RO2 + NO is entirely dependent on the branching ratio (alpha3) of the RO2 + NO reaction pathway. The smaller alpha3 is, the more important the RO + NO2 pathway is.
-- At 10 ppb of NO2 the RO + NO2 pathway accounts for 30 – 60\% of methyl nitrate production. At 100 ppb it accounts for 85 – 95\% of methyl nitrate production and almost 20\% of ethyl nitrate production.
- The rate of alkyl nitrate production in a polluted air mass is proportional to the OH concentration in that air mass.

>>>>

\subsection{Ozone, alkyl nitrates and carbon bonds}
The question that is always important for ozone-policies is the ozone production efficiency. Historically people noticed that with decreasing NOx emission the ozone production? concentration? drops, so the way of calculating the efficiency of ozone production due to drop in NOx emissions was proposed in the following way:

In addition to a traditional way of calculating ozone production efficiency, $\epsilon$, usually defined as the 
number of $O_3$ molecules produced per molecule of $NO_x$ consumed (Jacob, 1999): 
\begin{equation}
\epsilon = \dfrac{P_{O_3}}{L_{NO_x}}
\end{equation}
where $P_{O_3}$ – the $O_3$ production rate, and $L_{NO_x}$ – the $NO_x$ loss rate, a novel method of calculation of $\epsilon$ is applied.

This formula is not very useful if we want to estimate the impact of AN chemistry on O3, but there is another method, proposed by Professor Matthew Evans (University of York) suggests that the ozone production efficiency can be estimated via calculation of the number of carbon bonds emitted \citep{Evans2014} (personal communication). For example, if a given set of reactions contains only methane, $CH_4$, as 'a carbon bond reservoir', it means that this system has 4 carbon bonds (4 C-H). If  ethene, $C_2H_4$, is considered the system has 5 bonds (4 C-H bonds and 1 C=C) \citep{Edwards2013}. Assuming that every carbon bond is broken chemically and every peroxy radical ($RO_2$) reacts with an NO the overall efficiency of carbon bonds conversion to ozone can be calculated as: 
\begin{equation}
\epsilon = \dfrac{P_{O_3}}{E_{C_{bons}}} = \chi*nu
\end{equation}
where $P_{O_3}$ – the O3 production rate (mol hr-1), $E_{C_{bons}}$  – the carbon bond initial abundance (mol hr-1), $\chi$ – ratio between rate of carbon bonds loss to make $RO_2$ and rate of all carbon bonds loss, $nu$ – ratio between rate of $RO_2$ loss in reaction with NO and rate of all $RO_2$ loss. In other words, chemical loss of 1 carbon bond produces 1 peroxy radical, and photolysis loss of 1 carbon bond produces 0 or 2 peroxy. In other words,  it means that essentially we need to calculate the overall number of peroxy radicals remained in the chemical system by the end of simulation.

The same kind of approach is used in Common Representative Intermediates (CRI) mechanism for VOC degradation \citep{Jenkin2002, Jenkin2008}.
A reduced mechanism describing ozone formation from the tropospheric degradation of methane and 115 emitted non-methane hydrocarbons and oxygenated volatile organic compounds has been developed, using the Master Chemical Mechanism version 3.1 (MCM v3.1) as a reference benchmark. The Common Representative Intermediates mechanism version 2 (CRI v2) has been built up on a compound-by-compound basis, with the performance of its chemistry optimised for each compound in turn by comparison with that of MCM v3.1, using a series of five-day box model simulations. The resultant mech- anism contains 1183 reactions of 434 chemical species, i.e., ca. 10\% of the number of reactions and species to degrade the same set of VOCs in MCM v3.1. Similarly to CRI v1 (Jenkin et al., 2002a), a key assumption in the CRI v2 construction methodology is that the potential for ozone formation from a given volatile organic compound (VOC) is related to the number of reactive (i.e., C-C and C-H) bonds it contains. This index allows a series of generic intermediates to be defined, with each being used as a 'common representative' for a large set of species possessing the same index, as formed in detailed mechanisms such as the MCM.
On this basis, the number of reactive (C-H and C-C)
bonds in a closed-shell intermediate can be used as an index to represent the number of NO-to-NO2 conversions which can potentially result from its subsequent OH-initiated and NOx-catalysed degradation.

@@

\subsubsection*{Continent}
\citep{Day2003}
We describe measurements of NO2, total peroxy nitrates (?PNs), total alkyl nitrates (?ANs), and HNO3 using thermal dissociation followed by laser-induced fluorescence detection of NO2 at three continental locations. At a rural site in California, measurements over a full annual cycle show that ?ANs are routinely 10–20\% of NOy. rural site, at a suburban site in California and an urban site in Houston, Texas [rural (UC-BFRS), suburban (Granite Bay) and urban (La Porte)]
SumANs a more broad term than usual.
?AN mixing ratios at the UC-BFRS exhibit a strong seasonal cycle with a minimum in winter and a maximum in summer. The seasonal transition is sharp. Concentrations of ?ANs rise rapidly from 150 ppt at the end of April to 500 ppt by early June.
We also show that ?ANs are a large part, if not all, of the ‘‘missing NOy’’.

@@

\subsubsection*{Lifetime}
\citep{Newland2013}
[Species within this group provide a transport reservoir for the short lived NOx compounds due to their relatively longer lifetimes. This allows NOx to be transported from polluted areas to remote background sites. Alkyl nitrates have been reported as making up only a few percent of the NOy ('odd' nitrogen) budget in urban and rural areas (Shepson et al., 1993; Thornberry et al., 2001), however due to their relatively long lifetimes and their continued production during transport of polluted air masses, as alkanes are converted to alkyl nitrates (e.g. Bertman et al., 1995), they can comprise up to tens of percent of the NOy budget in polar regions and in the remote marine troposphere (Talbot et al., 2000). Therefore alkyl nitrates could potentially be a substantial source of NOx in remote background regions contributing to tropospheric ozone formation in these regions.
The lifetimes of the alkyl nitrates vary with latitude and season due to the seasonality of the OH sink and photolysis. Clemitshaw et al. (1997) report a range of lifetimes with respect to OH and to photolysis. Photolysis lifetimes vary from about ten days for butyl and pentyl nitrate in the summer to 1 – 2 years in the winter at 65N with lifetimes also decreasing with altitude. The lifetimes due to reaction with OH (tauOH) at 65N range from tens of years in
winter to about 10 days in the summer. tauOH increases with altitude as the OH concentration decreases. At the equator tauOH is similar all year round, on the order of a few days.]

\citep{Chuck2002}
Use of the mean sea-to-air fluxes for 0 to 40S (Table 1) and measured atmospheric concentrations results in a lifetime of be- tween 4.5 and 25 days for MeONO2 and between 5 and 10 days for EtONO2 (22). Literature estimates of the atmospheric life- time of MeONO2, based on known atmospheric destruction processes, are poorly con- strained, ranging from 6 to 29 days (23, 24). For EtONO2
, the atmospheric lifetime with respect to photodissociation at the equator and at 40S has been estimated to be 7 and 12 days, respectively, for October (25).

@@

\subsubsection{Comparison methodology}
\citep{Sommariva2008}
[The ratios between the alkyl nitrates and their
main alkane precursors can also be plotted against one another, for example against 2-butylnitrate/ n-butane as in Fig. 4. In this way the modelled and the measured concentrations can be compared without taking into account their variation with time. A similar approach was used in previous studies to investigate the sources of alkyl nitrates and the photochemical processing of polluted air masses (e.g., Bertman et al., 1995; Flocke et al., 1998; Roberts et al., 1998; Stroud et al., 2001; Simpson et al., 2003; Reeves et al., 2007). These studies compared plots of the measured ratios against 2-butylnitrate/n-butane to the ratios calculated with a simple sequential reaction model which assumed a single precursors for each alkyl nitrate. Usually, the underestimation of the ratios for short- chain compounds (propylnitrates/propane, ethylni- trate/ethane) was interpreted as an indication of photochemical sources other than the main alkane precursor.]

\citep{Day2003}
We also show that sumANs are a large part, if not all, of the 'missing NOy' reported in many prior experiments and emphasize that the ratios of sumANs/NOz and of O3/sumANs are especially useful for evaluating chemical models and comparing observations at different sites.

@@

\citep{Neu2008}
Observations have revealed widespread emissions of alkyl nitrates from the tropical and Southern Oceans. We present the first chemical transport model simulations to examine the global impact of these emissions. Matching observed atmospheric abundances, we derive a total oceanic flux of methyl nitrate (MeONO2) and ethyl nitrate (EtONO2) equivalent to 0.35 Tg of N per year, which contributes as much as 1 DU to the tropospheric ozone column in the Western Pacific and is responsible for about 3\% of the global oxidative capacity of the troposphere.
Atmospheric observations indicate that low molecular weight (C1-C4) alkyl nitrates are ubiquitous over the tropical and Southern Oceans [Atlas et al., 1993; Blake et al., 1999, 2003] and can be a significant component (20–80\%) of the total reactive nitrogen in remote marine air [Jones et al., 1999; Talbot et al., 2000]. Recent ship-borne measurements show supersaturation of alkyl nitrates in tropical surface waters, confirming an oceanic source [Chuck et al., 2002; Moore and Blough, 2002; Dahl et al., 2005]. This work presents the first 3-D chemical transport model (CTM) simulations to examine the role of oceanic alkyl nitrates in global chemistry, using observational constraints.
In this study, we incorporate MeONO2 and EtONO2 in the chemical mechanism of the UCI CTM [Carver et al., 1997] and compare their simulated abundances to observations to derive new estimates of their oceanic fluxes. We assess the relative importance of oceanic alkyl nitrates as a natural source of NOx in remote regions by evaluating their impact on tropospheric ozone (O3) and the oxidative ca- pacity of the atmosphere.
The UCI CTM [Wild and Prather, 2000; Hsu et al.,2005] includes 30 chemical species in the troposphere.
We define a control run with no alkyl nitrate emis-
sions and compare it to two simulations in which we impose a constant, spatially uniform flux of MeONO2 and EtONO2 from: 1) the tropical oceans (10?S–10?N) and 2) the Southern Ocean (south of 45?S).
Results: impact on NOx and O3 chemistry
Oceanic alkyl nitrates cause measurable increases in
both NOx and O3 in marine air. Their largest impact is seen over the equatorial Western Pacific (Figure 3). There is a persistent minimum in both NOx and O3 (Figure 3a) in this region, where the circulation isolates the marine boundary layer, resulting in sustained photochemical loss of O3 and a deep minimum throughout the tropospheric O3 column. At the same time, the circulation acts to accumulate alkyl nitrates in theWestern Pacific (see Figure 2), and they drive up boundary layer NOx abundances by as much as 250\%. This leads to O3 increases of up to 20\% (Figures 3b and 3f), with the largest increases occurring from May to October. From December to March (not shown), the relative increases in NOx and O3 are smaller, but are still concen- trated in the Western Pacific even though the alkyl nitrates are spread more evenly across the Pacific during these months. The geographic pattern of the alkyl nitrate influence persists into the middle and upper troposphere (Figures 3c-3d and 3g-3h), with reduced relative impact because of the larger lightning NOx source. However, the largest zonal mean increases in mixing ratio, 3–8 ppt for NOx and 1 ppb for O3, occur in the tropical upper and middle troposphere, respectively (Figure 4).
Pacific for example, ignore Southern ocean.
The O3 produced by NOx from alkyl nitrates fills in
the minimum in the tropospheric O3 column over the Western Pacific, adding more than 1 DU, with a maximum in October (Figure 3e). The global average increase in tropospheric O3 column varies with season from 0.35 to 0.42 DU.!!!!
WOW!!!
Oceanic alkyl nitrates are necessary to accurately
model NOx in the tropical remote marine environment, since they are responsible for 15–20\% of the O3 in the Western Pacific. The coincident location of the maximum alkyl nitrate impact and the deep minimum in O3 over the Western Pacific is a fascinating consequence of the convergence gence of the trade winds, but will make it difficult to detect an alkyl nitrate signature in O3. [22] At current levels, alkyl nitrates have a modest climate impact. From a simple scaling [Wild et al., 2001], the global increase in tropospheric column O3 will have a radiative forcing of about 16 mW/m2. In steady state, this will be almost entirely offset by the reduced CH4 lifetime, which produces a roughly equal negative radiative forcing. Surprisingly, a comparison to Wild et al. [2001] shows that on a per-molecule basis, alkyl nitrates have the largest impact on O3 and CH4 of any NOx source analyzed to date. This seems to be a consequence of their long lifetime combined with efficient transport to the tropical upper troposphere. With a 2.4\% effect on CH4 lifetime, a 5-fold increase in alkyl nitrate emissions during glacial periods would account for roughly 1/3 of the 300 ppb drop in CH4 seen between interglacials and glacials.

\citep{Atherton1989}
The transport from urban areas can lead to alkyl nitrate concentrations of 4-83 pptv after 120- 240 hours. These levels are far higher than those possible from marine in-situ production. Additionally, alkyl nitrates may be more abundant than PAN in remote marine environments.
Model: box
Chemistry: explicit chemical reaction mechanism
[Carter et al., 1986. A variety of alkyl nitrates, including methyl-, butyl-, pentyl-, hexyl-, heptyl-, octyl-, and nonyl-nitrates are allowed to form from fourteen different alkyl peroxy radicals. + ocean emission sources fluxes of c
for ethene, ethane, propene, and propane, NO, NO2
Initial conditions: clear sky conditions, a quantum yield of 1, solar zenith angle of 45 degrees, and total ozone column of 8*10 cm-2. Daytime+nigthtime, invertion height, dilution. 298 K. No deposition.
Scenarios: Two, MLOW and MHIGH, represent the marine in-situ production of alkyl nitrates. The total hydrocarbon concentration was 17.6 ppbC. MLOW had NO = 4 pptv and NO2 = 4.6 pptv based on measurements over the Pacific Ocean. MHIGH had NO = 46 pptv, NO2 = 54 pptv, and PAN = 50 pptv. The two other scenarios, UHIGH and ULOW, simulate the transport of species from urban to remote environments. UHIGH contained 2150 ppbC and 140 ppbv NOx, and ULOW contained 270 ppbC and 18 ppbv NOx, giving each a HC/NOx ratio of 15.
Simulation time: 12 hours MLOW and MHIGH, 240 hours UHIGH and ULOW.
Results:
- The total alkyl nitrate concentration present after 12 hours is less than 0.4 pptv for MHIGH and less than 0.1 pptv for MLOW. These levels are much lower than observations, indicating that marine in-situ production can account for only a fraction of the total marine alkyl nitrate concentration.
- IHIGH The total alkyl nitrate concentration peaks during the first 6 hours and then drops to 3.1 ppbv after 12 hours. This value is com- parable with the 2 ppbv alkyl nitrate concentration de- termined in Los Angeles [Hanst et al., 1982]
- The model predicts alkyl nitrate concentrations of 21-83 pptv for UHIGH and 4-12 pptv for ULOW for the hours 120-240.
- General: The model predictions of 12-83 pptv (af- ter 120 hours) and 4-21 pptv (after 240 hours) are of the same order as observations.

\citep{Sommariva2008}
Validation: typical urban plume in the North-East of the United States, measurements taken during three flights of the NOAA WP-3D aircraft, which sampled a plume from the New York City area during the New England Air Quality Study (NEAQS) 2004 campaign
- Seven alkyl nitrates (ethyl- nitrate, i-propylnitrate, n-propylnitrate, 2-butylni- trate, 2-pentylnitrate, 3-pentylnitrate, 3-methyl-2- butylnitrate) were measured during these flights together with a full suite of VOCs and other chemical parameters, such as CO, O3,NOx, HNO3 and PANs (Neuman et al., 2006).
- The model results have been qualitatively compared to measurements of alkyl nitrates taken on 20–22 July 2004, when the NOAA WP-3D aircraft made three flights aimed to inter- cept the New York City plume.
Model: subset! of MCM v3.1 (The MCM is one of the most detailed chemical mechanism for tropospheric chemistry available)
- to study the formation processes and photochemical evolution of alkyl nitrates
Chemistry: oxidation mechanisms of 64 VOCs (50 hydrocar-
bons and 14 oxygenates) plus CH4 and CO, The inorganic chemistry mechanism was taken from the IUPAC recommendation (Atkinson et al., 2003). A dimethyl sulphide (DMS) mechanism (Sommariva et al., 2004) was also included. The photolysis rates were calculated using the MCM parametrization (Saunders et al., 2003) for end of July 2004 at the latitude and longitude of Boston, MA. only gas-phase!!! Measurements made during NEAQS 2004 showed that there was very little removal of, for example,HNO3 by dry deposition in plumes decoupled from the marine boundary layer during transport from New York City (NYC) to the North Atlantic (Neuman et al., 2006). Therefore, removal by dry deposition was neglected in the model.
Initialisation: measurements taken in the Boston metropolitan area during the 2002 and 2004 NEAQS campaigns. The initial concentrations of VOCs were determined using emission ratios; CH4 and O3 were set to typical values for urban conditions (1800 and 90 ppb) (Table 1). The model has been initialized using measurements and VOCs emission ratios from the 2002 and 2004 New England Air Quality Studies.
Simulation time: 5 days.
Scenarios: 2: A (higher OH, good with obs, reference), B
Time: 20-22 July.
Results:
- While long-chain (C5) alkyl nitrates are produced for 90\% or more from the oxidation of a single parent alkane, short-chain <C4 alkyl nitrates can be formed from several precursors. The relative importance of each production route has been quantitatively determined thanks to the high level of chemical detail provided by the MCM. These secondary routes
to the formation of alkyl nitrates include the oxidation of longer-chain alkanes and oxygenated intermediates, like carbonyls, peroxides and carboxylic acids.
- The best agreement is in the case of 3-methyl-2-butylnitrate and the discrepancy between the modelled and the measured ratios clearly increases with decreasing carbon number .
- RH/AN, Except in the case of ethylnitrate, the model can reproduce the measured rate of increase in the alkyl nitrate to alkane ratios fairly well, which suggests that the kinetic parameters used in the MCMmodel are overall correct.
- By considering together Figs. 3 and 4 it can be reasonably concluded that the chemistry of alkyl nitrates is well represented in the MCM model.
- Using the explicit description of the chemistry provided by the MCM, it has been possible to determine quantitatively the photochemical formation and the evolution of alkyl nitrates and the relative importance of their different formation pathways. The results are summarized in Table 2. Large alkyl nitrates are produced almost exclusively from the reaction of a single parent
alkane. Smaller <=C4 alkyl nitrates also have other significant precursors, like longer-chain alkanes and alkenes. For n-propylnitrate and ethylnitrate, pro- panal and butanal, which are largely emitted, are important sources within the first 24 h. For all alkyl nitrates, the role of organic peroxides and car- boxylic acids formed by the reactions of peroxy radicals with HO2 at low [NOx] increases with time. These species contribute to sustain the production of alkyl nitrates far from the emission points of the parent VOCs. The relative importance of the different pathways
and how they change with time are dependent on the initial conditions of the model. In a scenario with lower O3 and CO and higher NOx the
importance of the direct routes (alkane ?OH) is typically lower during the first two–three days, and the contributions of peroxides and carboxylic acids become significant later and are on average less important. These results suggest that secondary chemistry in the formation of the alkyl nitrates, particularly oC4, cannot be neglected and should be included in the chemical models, due to its potential impact on ozone-controlling strategies.

\citep{Newland2013}
Model1: 2-D atmospheric model
- to study the formation, loss and transport of the alkyl nitrates to the Arctic
- to derive northern hemisphere emissions of the six alkanes.
Chemistry:
- (with fixed NOx and OH field taken from full chemistry run with model by Hough and Johnson (1990))
Validation of ANs: comparison with measurements of alkyl nitrates in the Arctic.
Performance: recreate both the measured seasonality of the alkyl nitrate, and the peak mixing ratio, reasonably well.
Simulation time: 6 days?
Model2: box
Chemistry:
- alkanesmethane, ethane, propane, butane and pentane and ANs explicitly described (RO+NO2 too)
Model3: firn model
- to derive atmospheric histories of six alkanes and six alkyl nitrates from measurements in firn air from the NEEM site, Greenland (methyl, ethyl, 2-propyl, 2-butyl, 2+3-pentyl and 3- methyl-2-butyl nitrate), 1950-2008.
Results:
- There are two ways in which NOx mixing ratios can limit alkyl nitrate production: either through air masses being NO limited with respect to the formation of a given alkyl nitrate or by the NOx in an air mass being 'used up' (track back to Reeves)
-- Alkyl nitrate production becomes NOx saturated above a given NO mixing ratio. This limit increases with increasing carbon number.
-- The time taken for an air mass to 'use up' all the initial NOx is dependent mainly on the initial NOx mixing ratio.
- At high [NO2] (> 27 ppb), methyl nitrate production is dominated by the reaction of the alkoxy radical with NO2
- The regions that contribute most to modeled alkane and alkyl nitrate mixing ratios in the Arctic are Europe and Russia
- The importance of the RO + NO2 pathway compared to RO2 + NO is entirely dependent on the branching ratio (alpha3) of the RO2 + NO reaction pathway. The smaller alpha3 is, the more important the RO + NO2 pathway is.
-- At 10 ppb of NO2 the RO + NO2 pathway accounts for 30 – 60\% of methyl nitrate production. At 100 ppb it accounts for 85 – 95\% of methyl nitrate production and almost 20\% of ethyl nitrate production.
- The rate of alkyl nitrate production in a polluted air mass is proportional to the OH concentration in that air mass.
